---
title: "Betalain_quantification"
author: "twinkle1"
date: "2023-11-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(reshape2)
library(patchwork)
library(agricolae)
library(ggpubr)
library(cowplot)
library(ggbeeswarm)
```

Create output directory for plots

```{bash}
mkdir ../../plots/betalain_quantification
```


Read in measured data

```{r}
# read in data
betalain_quantification <- read_excel(path = "../../raw_data/betalain_quantification/Betalain_quantification_summary.xlsx",
                                      sheet = "Photometric_quantification")

# normalize to the same input weight
#betalain_quantification <- betalain_quantification %>%
#  mutate(A536 = ifelse(tissue == "leaf", A536 * (0.0167/0.037), A536),
#         A480 = ifelse(tissue == "leaf", A480 * (0.0167/0.037), A480))

# correct for betacyanin absorption at 480 nm
betalain_quantification <- betalain_quantification %>%
  mutate(A480 = A480 - (A536/3.1))



# calculate the mass of betalains in our samples
betalain_quantification <- betalain_quantification %>%
  mutate(cBetacyanin = (A536/60000)*2, # dilution of factor 2 through added MetOH
         nBetacyanin = cBetacyanin * 0.000275,
         mBetacyanin = nBetacyanin * 726.6, # molar mass amaranthin
         factor_to_1g = 1/fresh_weight,
         mBetacyanin_in_1g = mBetacyanin * factor_to_1g,
         mBetacyanin_in_1g_in_mg = mBetacyanin_in_1g * 1000) %>%
  mutate(mBetacyanin_in_1g_in_mg = mBetacyanin_in_1g_in_mg * cuvette_dilution) # correct for dilution in cuvette

head(betalain_quantification)

betalain_quantification <- betalain_quantification %>%
  mutate(cBetaxanthin = (A480/48000)*2,
         nBetaxanthin = cBetaxanthin * 0.000275,
         mBetaxanthin = nBetaxanthin * 324.333, # vulgaxanthin IV
         factor_to_1g = 1/fresh_weight,
         mBetaxanthin_in_1g = mBetaxanthin * factor_to_1g,
         mBetaxanthin_in_1g_in_mg = mBetaxanthin_in_1g * 1000) %>%
  mutate(mBetaxanthin_in_1g_in_mg = mBetaxanthin_in_1g_in_mg * cuvette_dilution) %>%
  select(tube_nr, line, batch, individual, accession, tissue, mBetacyanin_in_1g_in_mg, mBetaxanthin_in_1g_in_mg)

# exclude sample due to irregular betacyanin content
betalain_quantification <- betalain_quantification %>%
  filter(tube_nr != 65)


# melt dataframe
melted_quant <- melt(betalain_quantification,
                     id.vars = c("tube_nr", "line", "batch", "individual", "accession", "tissue"))
colnames(melted_quant)[7:8] <- c("metabolite", "content")

# add unique identifier to each individual of different lines from some PI number
melted_quant <- melted_quant %>%
  mutate(uniq_ind = paste0(line, "_", batch, "_", individual),
         accession = factor(accession, levels = c("PI 576485","PI 538323","PI 686465")),
         tissue = as.factor(tissue),
         batch = as.factor(batch))

# mean betalain content
mean_expression <- melted_quant %>%
  group_by(accession, tissue, metabolite) %>%
  summarise(mean_content = mean(content))
```

Conduct statistical analysis

```{r}
# conduct anova between BSA parents
betalains_leaf <- betalain_quantification %>%
  filter(tissue == "leaf")
betalains_flower <- betalain_quantification %>%
  filter(tissue == "flower")

# anova for leaf betacyanin content
leaf_bc_aov <- aov(mBetacyanin_in_1g_in_mg ~ accession * batch, data = betalains_leaf)
summary(leaf_bc_aov)
# tukey post hoc test
leaf_bc_post_hoc <- HSD.test(leaf_bc_aov, trt = "accession")
leaf_bc_post_hoc
# extract groups for leaf
leaf_bc_groups <- leaf_bc_post_hoc$groups
leaf_bc_groups$factors <- factor(rownames(leaf_bc_groups), levels = c("PI 576485","PI 538323","PI 686465"))


# anova for leaf betaxanthin content
leaf_bx_aov <- aov(mBetaxanthin_in_1g_in_mg ~ accession * batch, data = betalains_leaf)
summary(leaf_bx_aov)
# tukey post hoc test
#leaf_bx_post_hoc <- HSD.test(leaf_bx_aov, trt = "batch")
#leaf_bx_post_hoc
# extract groups for leaf
#leaf_bx_groups <- leaf_bx_post_hoc$groups
#leaf_bx_groups$factors <- as.factor(rownames(leaf_bx_groups))


# anova for flower betacyanin content
flower_bc_aov <- aov(mBetacyanin_in_1g_in_mg ~ accession * batch, data = betalains_flower)
summary(flower_bc_aov)
# tukey post hoc test
flower_bc_post_hoc <- HSD.test(flower_bc_aov, trt = "accession")
flower_bc_post_hoc
# extract groups for flower
flower_bc_groups <- flower_bc_post_hoc$groups
flower_bc_groups$factors <- factor(rownames(flower_bc_groups), levels = c("PI 576485","PI 538323","PI 686465"))


# anova for flower betaxanthin content
flower_bx_aov <- aov(mBetaxanthin_in_1g_in_mg ~ accession * batch, data = betalains_flower)
summary(flower_bx_aov)
# tukey post hoc test
flower_bx_post_hoc <- HSD.test(flower_bx_aov, trt = "accession")
flower_bx_post_hoc
# extract groups for flower
flower_bx_groups <- flower_bx_post_hoc$groups
flower_bx_groups$factors <- factor(rownames(flower_bx_groups), levels = c("PI 576485","PI 538323","PI 686465"))
```

```{r}
# from: https://stackoverflow.com/questions/54672468/ggplot2-how-to-nudge-the-position-of-points-in-geom-beeswarm
position_nudge_any <- function(x = 0, y = 0, position) {
  ggproto(NULL, PositionNudgeAny,
          nudge = ggplot2::position_nudge(x, y),
          position = position
  )
}


#' Internal class doing the actual nudging on top of the other operation
#' @keywords internal
PositionNudgeAny <- ggplot2::ggproto("PositionNudgeAny", ggplot2::Position,
  nudge = NULL,
  nudge_params = NULL,
  position = NULL,
  position_params = NULL,

  setup_params = function(self, data) {
   list(nudge = self$nudge,
        nudge_params = self$nudge$setup_params(data),
        position = self$position,
        position_params = self$position$setup_params(data))
  },

  setup_data = function(self, data, params) {
   data <- params$position$setup_data(data, params$position_params)
   params$nudge$setup_data(data, params$nudge_params)
  },

  compute_layer = function(self, data, params, layout) {
   data <- params$position$compute_layer(data, params$position_params, layout)
   params$nudge$compute_layer(data, params$nudge_params, layout)
  }
)
```

Plot data

```{r}
# plot data
# plot betalain quantification from leaf
p_leaf <- ggplot(data = melted_quant %>% filter(tissue == "leaf")) +
  # geom_boxplot(aes(x = accession,
  #                y = content,
  #                fill = metabolite),
  #              outlier.shape = NA,
  #            color = "black",
  #            size = 0.4) +
  geom_point(data = melted_quant %>% filter(tissue == "leaf", metabolite == "mBetacyanin_in_1g_in_mg"),
             aes(x = accession, y = content),
             size = 1.8,
             color = "red3",
             position = position_nudge_any(x = -0.2,
                                           y = 0,
                                           position_beeswarm(priority = "random",
                                                             cex = 1.5))) +
  geom_point(data = melted_quant %>% filter(tissue == "leaf", metabolite == "mBetaxanthin_in_1g_in_mg"),
             aes(x = accession, y = content),
             color = "#F0B327",
             size = 1.8,
             position = position_nudge_any(x = 0.2,
                                           y = 0,
                                           position_beeswarm(priority = "random",
                                                             cex = 1.5))) +
  geom_segment(data = mean_expression %>% filter(tissue == "leaf", metabolite == "mBetacyanin_in_1g_in_mg"),
               aes(x = as.numeric(accession) - 0.35,
                   xend = as.numeric(accession) - 0.05,
                   y = mean_content,
                   yend = mean_content),
               linewidth = 0.9,
               color = "black") +
    geom_segment(data = mean_expression %>% filter(tissue == "leaf", metabolite == "mBetaxanthin_in_1g_in_mg"),
               aes(x = as.numeric(accession) + 0.05,
                   xend = as.numeric(accession) + 0.35,
                   y = mean_content,
                   yend = mean_content),
               linewidth = 0.9,
               color = "black") +
  geom_text(data = leaf_bc_groups, 
          aes(x = as.numeric(factors) - 0.19,
              y = 1.15,
              label = groups), 
          size=8, 
          inherit.aes = F,
          color = "red3") +  
  scale_color_manual(values = c( "red3","#F0B327"), labels = c("Betacyanins", "Betaxanthins")) +
  #scale_fill_manual(values = alpha(c("red3","#F0B327"), 0.2), labels = c("Betacyanins", "Betaxanthins")) +
  scale_fill_manual(values = c("red3","#F0B327"), labels = c("Betacyanins", "Betaxanthins")) +
  coord_cartesian(ylim = c(0, 1.2)) +
  labs(y = "Betalain content\n [mg/g]", color = "Wavelength") +
  theme_classic() +
    theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1)) +
        axis.text.x = element_blank()) +
  guides(color = guide_legend(override.aes = list(size = 4)))

p_leaf

# plot betalain quantification from flower
p_flower <- ggplot(data = melted_quant %>% filter(tissue == "flower")) +
  # geom_boxplot(aes(x = accession, 
  #                y = content, 
  #                fill = metabolite),
  #            color = "black",
  #            #color = "red4",
  #            size = 0.4) +
  geom_point(data = melted_quant %>% filter(tissue == "flower", metabolite == "mBetacyanin_in_1g_in_mg"),
             aes(x = accession, y = content),
             size = 1.8,
             color = "red3",
             position = position_nudge_any(x = -0.2,
                                           y = 0,
                                           position_beeswarm(priority = "random",
                                                             cex = 1.2))) +
  geom_point(data = melted_quant %>% filter(tissue == "flower", metabolite == "mBetaxanthin_in_1g_in_mg"),
             aes(x = accession, y = content),
             color = "#F0B327",
             size = 1.8,
             position = position_nudge_any(x = 0.2,
                                           y = 0,
                                           position_beeswarm(priority = "random",
                                                             cex = 1.2))) +
  geom_segment(data = mean_expression %>% filter(tissue == "flower", metabolite == "mBetacyanin_in_1g_in_mg"),
               aes(x = as.numeric(accession) - 0.35,
                   xend = as.numeric(accession) - 0.05,
                   y = mean_content,
                   yend = mean_content),
               linewidth = 0.9,
               color = "black") +
    geom_segment(data = mean_expression %>% filter(tissue == "flower", metabolite == "mBetaxanthin_in_1g_in_mg"),
               aes(x = as.numeric(accession) + 0.05,
                   xend = as.numeric(accession) + 0.35,
                   y = mean_content,
                   yend = mean_content),
               linewidth = 0.9,
               color = "black") +
  geom_text(data = flower_bc_groups, 
          aes(x = as.numeric(factors) - 0.19,
              y = 1.15,
              label = groups), 
          size=8, 
          inherit.aes = F,
          color = "red3") +  
  geom_text(data = flower_bx_groups, 
          aes(x = as.numeric(factors) + 0.19,
              y = 1.15,
              label = groups), 
          size=8, 
          inherit.aes = F,
          color = "#F0B327") +    
  scale_fill_manual(values = c( "red3","#F0B327"), labels = c("Betacyanins", "Betaxanthins")) +
  coord_cartesian(ylim = c(0, 1.2)) +
  labs(y = "Betalain content\n [mg/g]", color = "Wavelength") +
  theme_classic() +
  theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        #axis.title.y = element_text(size = 17),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1)) +
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank())

p_flower

patchplot_betalains <- p_flower / p_leaf +
  plot_annotation(tag_levels = "A") +
  plot_layout(guides = "collect", heights = c(1,1)) &
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        #plot.margin = unit(c(0,0,0,0.1), "cm"),
        plot.tag = element_text(vjust = 2, size = 24))

patchplot_betalains

ggsave(filename = "../../plots/betalain_quantification/betalain_quant_photometric_content.png",
       plot = patchplot_betalains,
       bg = "white",
       dpi = 450,
       width = 7,
       height = 9)
```


HPLC quantification

```{r}
# hplc quantification
hplc_quantification <- read_excel(path = "../../raw_data/betalain_quantification/Betalain_quantification_summary.xlsx",
                                      sheet = "LC-MS_quantification")

# get blank measurements
blank_amaranthin <- hplc_quantification[grep("Blank", hplc_quantification$sample_name), "Amaranthin_ratio"]
blank_betanin <- hplc_quantification[grep("Blank", hplc_quantification$sample_name), "Betanin_ratio"] # not detected
blank_betalamic_acid <- hplc_quantification[grep("Blank", hplc_quantification$sample_name), "Betalamic_acid_ratio"] # not detected
blank_vulgaxanthin_IV <- hplc_quantification[grep("Blank", hplc_quantification$sample_name), "Vulgaxanthin_IV_ratio"] # not detected

# join with other table
hplc_quant_joined <- inner_join(hplc_quantification, betalain_quantification) %>%
  mutate_all(~replace(., is.na(.), 0)) %>%
  mutate(uniq_ind = paste0(accession, "_", line, "_", individual))

# adjust measurements for different input weights
hplc_quant_joined <- hplc_quant_joined %>%
  mutate(Amaranthin_ratio = ifelse(tissue == "leaf", Amaranthin_ratio * (0.0167/0.037), Amaranthin_ratio),
         Betanin_ratio = ifelse(tissue == "leaf", Betanin_ratio * (0.0167/0.037), Betanin_ratio),
         Betalamic_ratio = ifelse(tissue == "leaf", Betalamic_acid_ratio * (0.0167/0.037), Betalamic_acid_ratio),
         Vulgaxanthin_IV_ratio = ifelse(tissue == "leaf", Vulgaxanthin_IV_ratio * (0.0167/0.037), Vulgaxanthin_IV_ratio))

# subtract blank
hplc_quant_normalised <- hplc_quant_joined %>% # replace NA with 0
  mutate(Amaranthin_ratio = Amaranthin_ratio - blank_amaranthin[[1]],
         accession = factor(accession, levels = c("PI 576485","PI 538323","PI 686465"))) # subtract blank from actual measurements

# replace negative values with 0
hplc_quant_normalised[hplc_quant_normalised < 0] <- 0


# AUC means
hplc_AUC_means <- hplc_quant_normalised %>%
  group_by(tissue, accession) %>%
  summarise(amaranthin_mean = mean(Amaranthin_ratio),
            betanin_mean = mean(Betanin_ratio),
            betalamic_acid_mean = mean(Betalamic_acid_ratio),
            vulgaxanthin_IV_mean = mean(Vulgaxanthin_IV_ratio)) %>%
  unique()


```


Statistical analysis

```{r}
# conduct anova between BSA parents
hplc_leaf <- hplc_quant_normalised %>%
  filter(tissue == "leaf")
hplc_flower <- hplc_quant_normalised %>%
  filter(tissue == "flower")


anova_and_tukey <- function(data, column){
  out_list <- list()
  # anova
  aov_out <- aov(reformulate("accession", response = column), data = data)
  out_list[[1]] <- aov_out
  # tukey post hoc
  post_hoc_out <- HSD.test(aov_out, trt = "accession")
  out_list[[2]] <- post_hoc_out
  # extract groups
  groups_out <- post_hoc_out$groups
  groups_out$factors <- factor(rownames(groups_out), levels = c("PI 576485","PI 538323","PI 686465"))
  out_list[[3]] <- groups_out
  return(out_list)
}

# amaranthin leaf
amaranthin_leaf <- anova_and_tukey(data = hplc_leaf, column = "Amaranthin_ratio")
summary(amaranthin_leaf[[1]])
amaranthin_leaf[2]

# amaranthin flower
amaranthin_flower <- anova_and_tukey(data = hplc_flower, column = "Amaranthin_ratio")
summary(amaranthin_flower[[1]])
amaranthin_flower[2]

# betanin leaf
betanin_leaf <- anova_and_tukey(data = hplc_leaf, column = "Betanin_ratio")
summary(betanin_leaf[[1]])
betanin_leaf[2]

# betanin flower
betanin_flower <- anova_and_tukey(data = hplc_flower, column = "Betanin_ratio")
summary(betanin_flower[[1]])
betanin_flower[2]

# betalamic acid leaf
ba_leaf <- anova_and_tukey(data = hplc_leaf, column = "Betalamic_acid_ratio")
summary(ba_leaf[[1]])
ba_leaf[2]

# betalamic acid flower
ba_flower <- anova_and_tukey(data = hplc_flower, column = "Betalamic_acid_ratio")
summary(ba_flower[[1]])
ba_flower[2]

# vulgaxanthin leaf
vulgaxanthin_leaf <- anova_and_tukey(data = hplc_leaf, column = "Vulgaxanthin_IV_ratio")
summary(vulgaxanthin_leaf[[1]]) # no significant difference

# vulgaxanthin flower
vulgaxanthin_flower <- anova_and_tukey(data = hplc_flower, column = "Vulgaxanthin_IV_ratio")
summary(vulgaxanthin_flower[[1]]) # no significant difference
```

Plot results amaranthin

```{r}
# plot
pa_leaf <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "leaf")) +
  # geom_point(aes(x = accession, 
  #                y = Amaranthin_ratio, 
  #                group = uniq_ind),
  #            color = "red3",
  #            position = position_dodge(width = 0.75),
  #            size = 2.8) +
  geom_beeswarm(aes(x = accession, 
                 y = Amaranthin_ratio, 
                 group = uniq_ind),
             color = "red3",
             cex = 1.5,
             size = 2.8) +
  geom_segment(data = hplc_AUC_means %>% filter(tissue == "leaf"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = amaranthin_mean, yend = amaranthin_mean),
                color = "black",
                linewidth = 0.9) +
  geom_text(data = amaranthin_leaf[[3]],
          aes(x = as.numeric(factors),
              y = 280,
              label = groups),
          size=8,
          inherit.aes = F,
          color = "red3") +
  coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Amaranthin\n relative area", color = "Wavelength") +
  scale_color_manual(values = c("red3", "red3","black","#F0B327")) +
  theme_classic() +
  theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1),
        axis.text.x = element_blank(),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))



pa_flower <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "flower")) +
  # geom_point(aes(x = accession, 
  #                y = Amaranthin_ratio, 
  #                group = uniq_ind),
  #            color = "red3",
  #            position = position_dodge(width = 0.75),
  #            size = 2.8) +
  geom_beeswarm(aes(x = accession, 
                 y = Amaranthin_ratio, 
                 group = uniq_ind),
             color = "red3",
             cex = 1.5,
             size = 2.8) +
  geom_segment(data = hplc_AUC_means %>% filter(tissue == "flower"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = amaranthin_mean, yend = amaranthin_mean),
                color = "black",
                linewidth = 0.9) +
  geom_text(data = amaranthin_flower[[3]],
          aes(x = as.numeric(factors),
              y = 280,
              label = groups),
          size=8,
          inherit.aes = F,
          color = "red3") +
  coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Amaranthin\n relative area", color = "Wavelength") +
  scale_color_manual(values = c("red3", "red3","black","#F0B327")) +
  theme_classic() +
    theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        #axis.title.y = element_text(size = 17),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1),
        axis.text.x = element_blank(),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))

```

Plot results betanin

```{r}
# plot
pb_leaf <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "leaf")) +
    geom_beeswarm(aes(x = accession, 
                 y = Betanin_ratio, 
                 group = uniq_ind),
             color = "red3",
             cex = 1.5,
             size = 2.8) +
  geom_segment(data = hplc_AUC_means %>% filter(tissue == "leaf"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = betanin_mean, yend = betanin_mean),
                color = "black",
                linewidth = 0.9) +
  geom_text(data = betanin_leaf[[3]],
          aes(x = as.numeric(factors),
              y = 280,
              label = groups),
          size=8,
          inherit.aes = F,
          color = "red3") +
  coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Betanin\n relative area", color = "Wavelength") +
  scale_color_manual(values = c("red3", "red3","black","#F0B327")) +
  theme_classic() +
  theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1),
        axis.text.x = element_blank(),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))



pb_flower <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "flower")) +
    geom_beeswarm(aes(x = accession, 
                 y = Betanin_ratio, 
                 group = uniq_ind),
             color = "red3",
             cex = 1.5,
             size = 2.8) +
  geom_segment(data = hplc_AUC_means %>% filter(tissue == "flower"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = betanin_mean, yend = betanin_mean),
                color = "black",
                linewidth = 0.9) +
  geom_text(data = betanin_flower[[3]],
          aes(x = as.numeric(factors),
              y = 280,
              label = groups),
          size=8,
          inherit.aes = F,
          color = "red3") +
  coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Betanin\n relative area", color = "Wavelength") +
  scale_color_manual(values = c("red3", "red3","black","#F0B327")) +
  theme_classic() +
    theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        #axis.title.y = element_text(size = 17),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1),
        axis.text.x = element_blank(),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))
```

Plot results betalamic acid

```{r}
# plot
pba_leaf <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "leaf")) +
    geom_beeswarm(aes(x = accession, 
                 y = Betalamic_acid_ratio, 
                 group = uniq_ind),
             color = "grey40",
             cex = 1.5,
             size = 2.8) +
  geom_segment(data = hplc_AUC_means %>% filter(tissue == "leaf"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = betalamic_acid_mean, yend = betalamic_acid_mean),
                color = "black",
                linewidth = 0.9) +
  geom_text(data = ba_leaf[[3]],
          aes(x = as.numeric(factors),
              y = 280,
              label = groups),
          size=8,
          inherit.aes = F,
          color = "grey40") +
  coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Betalamic acid\n relative area", color = "Wavelength") +
  scale_color_manual(values = c("red3", "red3","black","#F0B327")) +
  theme_classic() +
  theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 14, color = "black"),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1),
        axis.text.x = element_blank(),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))



pba_flower <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "flower")) +
    geom_beeswarm(aes(x = accession, 
                 y = Betalamic_acid_ratio, 
                 group = uniq_ind),
             color = "grey40",
             cex = 1.5,
             size = 2.8) +
  geom_segment(data = hplc_AUC_means %>% filter(tissue == "flower"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = betalamic_acid_mean, yend = betalamic_acid_mean),
                color = "black",
                linewidth = 0.9) +
  geom_text(data = ba_flower[[3]],
          aes(x = as.numeric(factors),
              y = 280,
              label = groups),
          size=8,
          inherit.aes = F,
          color = "grey40") +
  coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Betalamic acid\n relative area", color = "Wavelength") +
  scale_color_manual(values = c("red3", "red3","black","#F0B327")) +
  theme_classic() +
    theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        #axis.title.y = element_text(size = 17),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        #axis.text.x = element_text(size = 13, angle = 45, vjust = 1, hjust=1),
        axis.text.x = element_blank(),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))
```

Plot results vulgaxanthin

```{r}
# plot
pv_leaf <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "leaf")) +
  geom_beeswarm(aes(x = accession, 
                 y = Vulgaxanthin_IV_ratio, 
                 group = uniq_ind),
             color = "#F0B327",
             method = "swarm",
             cex = 1.5,
             size = 2.8) +
   geom_segment(data = hplc_AUC_means %>% filter(tissue == "leaf"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = vulgaxanthin_IV_mean, yend = vulgaxanthin_IV_mean),
                color = "black",
                linewidth = 0.9) +
  ylim(c(0, 300)) +
  #coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Vulgaxanthin IV\n relative area", color = "Wavelength") +
  theme_classic() +
  scale_x_discrete(labels = c("PI 576485\n leaf","PI 538323\n leaf","PI 686465\n leaf")) +
  theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        axis.title.y = element_text(size = 15),
        axis.text.y = element_text(size = 14, color = "black"),
        axis.text.x = element_text(size = 15, color = "black"),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))



pv_flower <- ggplot(data = hplc_quant_normalised %>% filter(tissue == "flower")) +
  geom_beeswarm(aes(x = accession, 
                 y = Vulgaxanthin_IV_ratio, 
                 group = uniq_ind),
             color = "#F0B327",
             method = "swarm",
             cex = 1.5,
             size = 2.8) +
   geom_segment(data = hplc_AUC_means %>% filter(tissue == "flower"),
                aes(x = as.numeric(accession)-0.2, xend = as.numeric(accession)+0.2,
                    y = vulgaxanthin_IV_mean, yend = vulgaxanthin_IV_mean),
                color = "black",
                linewidth = 0.9) +
  ylim(c(0, 300)) +
  #coord_cartesian(ylim = c(0, 300)) +
  labs(y = "Vulgaxanthin IV\n relative area", color = "Wavelength") +
  theme_classic() +
  scale_x_discrete(labels = c("PI 576485\n flower","PI 538323\n flower","PI 686465\n flower")) +
    theme(text = element_text(size = 24),
        axis.title.x = element_blank(),
        #axis.title.y = element_text(size = 17),
        legend.title = element_blank(),
        plot.margin = unit(c(0,0,0,0), "cm"),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.text.x = element_text(size = 15, color = "black"),
        legend.position = "none") +
  guides(color = guide_legend(override.aes = list(size = 4)))
```



Overview over all quantified metabolites including photometric quantification

```{r}
overview_patchplot <- p_leaf + p_flower + pa_leaf + pa_flower + pb_leaf + pb_flower +
  pba_leaf + pba_flower + pv_leaf + pv_flower +
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(ncol = 2,
              byrow = T,
              guides = "collect") &
  theme(legend.position = "bottom",
        legend.direction = "vertical",
        plot.tag = element_text(vjust = 2, size = 17, face = "bold"))

ggsave(filename = "../../plots/betalain_quantification/overview_betalain_quantification.png",
       plot = overview_patchplot,
       bg = "white",
       dpi = 450,
       width = 10,
       height = 13)
```


Combine specific plots


```{r}
# combine for figure
photo_hplc_patchplot <- plot_spacer() / (p_leaf + p_flower) / (pa_leaf + pa_flower) / (pv_leaf + pv_flower) +
  plot_annotation(tag_levels = list(c("(b)","(c)","(d)","(e)","(f)","(g)"))) +
  plot_layout(widths = c(0.5, 0.5),
              guides = "collect") &
  theme(legend.position = "bottom",
        plot.tag = element_text(size = 17, face = "bold"))



ggsave(filename = "../../plots/betalain_quantification/quantification_without_picture.png",
       plot = photo_hplc_patchplot,
       bg = "white",
       dpi = 450,
       width = 10,
       height = 10)
```


Analyse HPLC results from amaranth roots:

```{r}
# hplc quantification
root_quant <- read_excel(path = "../../raw_data/betalain_quantification/Betalain_quantification_summary.xlsx",
                                      sheet = "transgenic_roots_LC-MS_quantification")
root_quant <- root_quant %>%
  mutate(sample_id = paste0("plate_", plate, "_", root_type)) %>%
  mutate(sample_id = factor(sample_id, levels = c("plate_3_white", "plate_3_red", "plate_4_white", "plate_4_red")))

# normalise for solvent volume and input weight
norm_factor <- (75/275) * (0.0167/0.0082) # initial solvent volume 75 instead of 275, normalised input weight 8.2 mg instead of 16.7 mg
root_quant_norm <- root_quant %>%
  mutate(Amaranthin_ratio = Amaranthin_ratio * norm_factor,
         plant_replicate = factor(plant_replicate))


# plot comparison
root_quant_plot <- ggplot(data = root_quant_norm) +
  geom_col(aes(x = plant_replicate,
               y = Amaranthin_ratio,
               fill = root_type),
           color = "black",
           position = position_dodge2(reverse = T)) +
  geom_text(aes(x = plant_replicate,
                y = Amaranthin_ratio,
                label = round(Amaranthin_ratio,2),
                group = root_type),
            position = position_dodge2(width = 0.9, reverse = T),
            vjust = -0.2,
            size = 5) +
  scale_x_discrete(labels = c("Ind 1", "Ind 2")) +
  coord_cartesian(ylim = c(0, 450)) +
  labs(y = "Amaranthin\n relative area", x = "") +
  scale_fill_manual(values = c("red3", "beige")) +
  theme_classic() +
  theme(text = element_text(size = 24),
        legend.position = "none")
root_quant_plot

ggsave(filename = "../../plots/betalain_quantification/root_quantification.png",
       plot = root_quant_plot,
       bg = "white",
       dpi = 450,
       width = 7,
       height = 5)

```


